---
title: "Diatom Data Compilation"
author: "Zachary M. Smith"
date: "9/23/2019"
output: html_document
---

# Purpose

Brian Duffy asked me to automate the process for appending diatom data to pass the data on to ITS for the data modernization project.

__These files still contain old site names that need to be updated.__

# Script

Import the necessary R-packages
```{r}
library(tidyverse)
```

Establish path way to data files.
```{r}
data.path <- file.path("L:",
                       "DOW",
                       "SMAS",
                       "projects",
                       "its_data_modernization",
                       "RawData",
                       "DiatomData")
```

Identify all of the XLSX files in the specified data path. Split the identified XLSX files into two vectors representing metric and deliverable files.
```{r}
all_files.vec <- list.files(data.path,
                            pattern = "xlsx",
                            recursive = TRUE)

metric_files.vec <- all_files.vec[grepl("Metric", all_files.vec)]
deliverables_files.vec <- all_files.vec[grepl("Deliverables", all_files.vec)]
```

Create a standard function for importing and combining multiple files of the same type.
```{r}
# file.i <- metric_files.vec[1]
append_files <- function(files.vec) {
  files.vec %>% 
  lapply(function(file.i) {
    sub.df <- readxl::read_xlsx(
      file.path(data.path,
                file.i),
      sheet = 1,
      col_types = "text"
    )
    
    # sub.df$origin <- file.i
    
    sub.df
  }) %>% 
  dplyr::bind_rows()
}
```

Import, append, and export the metric files to the "compiled" folder.
```{r}
metrics.df <- append_files(files.vec = metric_files.vec)

data.table::fwrite(metrics.df,
          file.path(data.path,
                    "compiled",
                    "diatom_community-metrics.csv"))
```

Import, append, and export the deverable files to the "compiled" folder.
```{r}
deliverables.df <- append_files(files.vec = deliverables_files.vec)

data.table::fwrite(metrics.df,
          file.path(data.path,
                    "compiled",
                    "diatom_deliverables-counts.csv"))
```

