---
title: "Fish Site ID Crosswalk"
author: "Zachary M. Smith"
date: "10/25/2019"
output: html_document
---

# Purpose

Crosswalk old site IDs with new site IDs.

# Script

Import the necessary R-packages
```{r}
library(tidyverse)
```

Establish path way to data files.
```{r}
data.path <- file.path("L:",
                       "DOW",
                       "SMAS",
                       "projects",
                       "its_data_modernization",
                       "RawData")
```

Import old site files table to be used as a cross walk.
```{r}
sites.df <- readxl::read_xlsx(
      file.path(data.path,
                "OldSiteFile.xlsx"),
      sheet = 1
    ) %>% 
  mutate(RIVMILE = as.numeric(RIVMILLE)) %>% 
  select(SITE_ID,
         RIVMILE,
         STATION,
         LOCATION,
         LATITUDE,
         LONGITUDE)

```

Import the most recent site IDs (basin, location, and rivermile).
```{r}
blr.df <- read_csv(
  file.path(data.path,
            "sites_field.csv")
) %>% 
  select(SITE_ID,
         SITE_LOC_ID)
```

Import the fish data.
```{r}
fish.df <- readxl::read_xlsx(
      file.path(data.path,
                "Fish",
                "FishData.xlsx"),
      sheet = 1
    )
```

The "LOCATION" colmun in `fish.df` contains some rows that actually represent "SITE_ID". `fish.df` is split into two data frames representing rows where "LOCATION" actually representes "LOCATION" strings and rows where "LOCATION" represents "SITE_ID". The two dataframes are joined with the `sites.df` dataframe based on common columns.
```{r}
fish_loc_char.df <- fish.df %>% 
  filter(is.na(readr::parse_integer(LOCATION))) %>% 
  left_join(sites.df, by = c("LOCATION",
                             "RIVMILE"))

fish_loc_int.df <- fish.df %>% 
  filter(!is.na(readr::parse_integer(LOCATION))) %>% 
  mutate(SITE_ID = as.numeric(LOCATION)) %>% 
  select(-LOCATION, -RIVMILE) %>% 
  left_join(sites.df, by = c( "SITE_ID")) 
```

The two data frames are appended together.
```{r}
fish_sites.df <- bind_rows(fish_loc_char.df, fish_loc_int.df)
```

`fish_sites.df`  is merged with the new site ID table. There was one instance where the "SITE_ID" in te fish tables
```{r}
fish_final.df <- left_join(fish_sites.df, blr.df, by = "SITE_ID")%>% 
  select(SITE_LOC_ID, everything()) %>% 
  mutate(SITE_LOC_ID = if_else(SITE_LOC_ID == "BIGC" & RIVMILE == 6.2,
                               "10-BIGC-6.3",
                               SITE_LOC_ID)) %>% 
  select(SITE_LOC_ID, !!!names(fish.df)) %>% 
  select(-BASIN, LOCATION, RIVMILE)
```

```{r}
readr::write_csv(fish_final.df,
                 file.path("L:",
                       "DOW",
                       "SMAS",
                       "projects",
                       "its_data_modernization",
                       "cleaned",
                       "fish",
                       glue::glue('{format(Sys.Date(), "%Y%m%d")}_fish.csv')),
                 na = "")
  
```

